Relative parabolic factorizations of Chevalley groups

### 2015-04-18 paper.tex
Added some text about representations

@@

\subsection{Relative elementary subgroup}
The following theorem is the main result of \cite{S}.
\begin{thm}\label{theorem:Stepanov}
Let $\Phi$ be a reduced irreducible root system and let $S$ be any parabolic subset of roots of $\Phi$ with special part $\Sigma_S$.
Then $\E(\Phi, R, I)$ is generated by the following two families of elements:
\begin{itemize}
 \item $x_{\alpha}(s)$, where $s\in I$, $\alpha\in\Phi$;
 \item $z_\alpha(s,\xi)$, where $s\in I$, $\xi\in R$, $\alpha\in\Sigma_S$. \end{itemize} \end{thm}
\begin{proof} See~\cite[Theorem~3.4]{S}. \end{proof}

@@

\section{Bass---Kolster decomposition}\label{sec:bass-kolster}
\input{bass-kolster}

>>>>

In this paper we work with irreducible representations of Chevalley groups corresponding to the highest weight
$\varpi_1$ (in the case of a classical $\Phi_l$) or $\varpi_l$ (in the case $\Phi_l=\rE_l$). 
The former are the natural vector representations of classical groups acting on free modules $V=R^n$ of dimension 
$n=l+1, 2l+1, 2l,2l$ for $\Phi=\rA_l,\rB_l,\rC_l,\rD_l$.
The latter representations have dimension $n=27, 56, 248$ for $l=6,7,8$.

For a representation $\pi$ of $\G(\Phi, R)$ on $V$ we denote by $\Lambda(\pi)$ the set of weights of $\pi$.
It will be convenient to use standard numbering for elements of $\Lambda(\pi)$ in the case when $\Phi$ is classical and $\pi$ is a vector representation (cp.~\cite[§1B]{St78}):
$$\begin{array}{cll}
  1,2,\ldots, l+1 & \text{in the case} & \Phi =\rA_l, \\
  1,2,\ldots l, 0, -l,\ldots, -2, -1 & \text{in the case} & \Phi =\rB_l, \\
  1,2,\ldots l, -l,\ldots, -2, -1 & \text{in the cases}   & \Phi =\rC_l, \rD_l. \\
\end{array}$$
In particular, we write $1$ instead of $\varpi_1$, $2$ instead of $\varpi_1-\alpha_1$ and so on.

\subsection{Elementary subgroup}

@@

The following theorem is the main result of \cite{S}.
\begin{thm}\label{theorem:Stepanov}
Let $\Phi$ be a reduced irreducible root system and let $S$ be any parabolic subset of roots of $\Phi$ with special part $\Sigma_S$.
Then $\E(\Phi, R, I)$ is generated by the following two families of elements:
\begin{itemize}
 \item $x_{\alpha}(s)$, where $s\in I$, $\alpha\in\Phi$;
 \item $z_\alpha(s,\xi)$, where $s\in I$, $\xi\in R$, $\alpha\in\Sigma_S$. \end{itemize} \end{thm}
\begin{proof} See~\cite[Theorem~3.4]{S}. \end{proof}

@@

\section{Relative Bass---Kolster decomposition}\label{sec:bass-kolster}
\input{bass-kolster}

### 2015-04-18 stability-conditions.tex
Added some text about representations

To show the converse take arbitrary row $(a_1,\ldots,a_n,d)\in{}^{n+1}\!R$ and consider the unimodular row $(a_1,\ldots,a_n,1,d)\in\Ums(n+2,R).$
There exist $c_1,\ldots,c_n$ such that \[(a'_1,\ldots,a'_n,1+td)=(a_1+c_1d,\ldots,a_n+c_nd,1+td)\] is unimodular for all $t\in R$.
Suppose that there exists a maximal ideal $\mathfrak{m}\trianglelefteq R$ such that $a'_1,\ldots,a'_n\in\mathfrak{m}$ but at least one of the elements $d$, $a_i$ is not in $\mathfrak{m}$.
Clearly, $d\notin\mathfrak{m}$ (otherwise $a_i=a'_i-c_id\in\mathfrak{m}$).
On the other hand, $1+td\notin\mathfrak{m}$, hence the inequality $1+\bar{t}\bar{d}\neq0$ in the residue field $R/\mathfrak{m}$ may not hold for arbitrary $\bar{t}$ (take e.g. $\bar{t}=-1/\bar{d}$).
This shows that no such $\mathfrak{m}$ may exist and, therefore, $\mathfrak{J}(a_1,\ldots,a_n,d)=\mathfrak{J}(a'_1,\ldots,a'_n)$.
\end{proof}

@@

>>>>

To show the converse take arbitrary row $(a_1,\ldots,a_n,d)\in{}^{n+1}\!R$ and consider the unimodular row $(a_1,\ldots,a_n,1,d)\in\Ums(n+2,R).$
There exist $c_1,\ldots,c_n$ such that \[(a'_1,\ldots,a'_n,1+td)=(a_1+c_1d,\ldots,a_n+c_nd,1+td)\] is unimodular for all $t\in R$.
Suppose that there exists a maximal ideal $\mathfrak{m}\trianglelefteq R$ such that $a'_1,\ldots,a'_n\in\mathfrak{m}$ but at least one of the elements $d$, $a_i$ is not in $\mathfrak{m}$.
Clearly, $d\notin\mathfrak{m}$ (otherwise $a_i=a'_i-c_id\in\mathfrak{m}$ contrary to the assumption).
On the other hand, $1+td\notin\mathfrak{m}$,
hence the inequality $1+\bar{t}\bar{d}\neq0$ in the residue field $R/\mathfrak{m}$ may not hold for arbitrary $\bar{t}$ 
(take e.g. $\bar{t}=-1/\bar{d}$).
This shows that no such $\mathfrak{m}$ may exist and, therefore, $\mathfrak{J}(a_1,\ldots,a_n,d)=\mathfrak{J}(a'_1,\ldots,a'_n)$.
\end{proof}

@@

The following lemma illustrates a typical application of stability conditions.
\begin{lemma} \label{lemma:uraction} 
Let $\Phi=\rA_\ell, \rD_\ell$. Denote by $\pi$ the vector representation of $\G(\Phi, R)$ on $V=R^n$.
Assume that one of the following conditions holds:
\begin{itemize}
 \item $\Phi=\rA_\ell$, $\Gamma=\{ k+1, \ldots, \ell+1\} \subset \Lambda(\pi)$ and $\sr(I)\leq k\leq \ell$;
 \item $\Phi=\rD_\ell$, $\Gamma=\{-l,\ldots, -2, -1\} \subset \Lambda(\pi)$ and $\asr(R, I)\leq \ell-1$. 
\end{itemize}
Then for any unimodular column $v\in \Ums(n, R, I)$ there exist $x\in \U(\Phi^+, I)$, $y\in \U(\Phi^-, I)$ such that $(yx \cdot v)_\lambda = 0$ for $\lambda\in \Gamma$.
\end{lemma}

%Let $\sr(I)\leq l-1$, then for any unimodular column $v\in R^l=V$ there exist
%matrices $x=\left(\begin{smallmatrix} e_{l-1}& * \\ 0 & 1 \end{smallmatrix}\right)$, $y=\left(\begin{smallmatrix} e_{l-1}& 0 \\ *& 1 \end{smallmatrix}\right)$
%such that the $l$-th cooordinate of the vector $yx\cdot v$ is zero.\end{lemma}

%Denote by $p$ the matrix which has $1$'s along its secondary diagonal and zeroes elsewhere. Call a matrix $a\in\M(l, R)$ {\it antipersymmetric} if $pa^tp = a$.
%\begin{lemma}\label{LambdaS_lemma} Assume that $\asr(R)\leq l-1$, then for any columns $u^+, u^-\in R^l$ such that $(u^+, u^-)^t\in R^{2l}$ is unimodular there exists
%an antipersymmetric matrix $a\in\M(l, R)$ such that $u^+ + au^- \in R^l$ is unimodular. \end{lemma}
%\begin{proof} The statement of the lemma is a special case of \cite[Th.~1.2]{BPT} applied to $R$ viewed as a form ring with trivial involution and zero form parameter. \end{proof}

%Denote by $H$ the hyperbolic embedding $H:\GL(l, R)\rightarrow\mathrm{O}(2l, R)$ of a general linear group into the even-dimensional orthogonal group,
%i.\,e. the map which sends $g$ to $\left(\begin{smallmatrix} g & 0 \\ 0 & p(g^{t})^{-1}p \end{smallmatrix}\right)$.
%\begin{lemma}\label{asrLemma} Assume that $\asr(R)\leq l-1$. Then for any unimodular column $v\in R^{2l}=V$ there exist orthogonal matrices 
%$x=\left(\begin{smallmatrix} e_l & * \\ 0 & e_l \end{smallmatrix}\right)$, $y=\left(\begin{smallmatrix} e_l & 0 \\ * & e_l \end{smallmatrix}\right)$
%such that $(yx\cdot v)_k=0$ for $k=-l,\ldots, -1.$ \end{lemma}
%\begin{proof}
% Applying Lemma~\ref{LambdaS_lemma} we obtain $x=\left(\begin{smallmatrix} e_l & a \\ 0 & e_l \end{smallmatrix}\right)$ such
% that the first $l$ coor\-dinates of $x\cdot v$ form a unimodular column of height $l$.

% The group $\E(l, R)$ acts transitively on the set of unimodular columns of height $l$, therefore
% we can choose an appropriate $g\in\E(l, R)$ such that $$H(g)x\cdot v= (1,0,\ldots, 0,*,\ldots *)^t,$$
% where the number of stars equals $l$.
% Choose $y'=\left(\begin{smallmatrix} e_l & 0 \\ * & e_l \end{smallmatrix}\right)$ such that $y'H(g)x\cdot v = e_1 = (1,0,\ldots, 0)^t$.
% Clearly $y=y'^{H(g)}$ is an orthogonal matrix of the required form and $$yx\cdot v = y'^{H(g)} x\cdot v = H(g)^{-1} e_1 = (*,\ldots, *,0,\ldots, 0)^t.$$
% \end{proof}
